ShrinkNets

### 2018-01-05 paper/paper.tex
Finish my writing pass on the paper

\srm{In this paper, we describe ... -- need to give a high level idea of
  solution, and its benefits.  Would be good to give your system/approach a
  name.}

@@

\srm{Describe our key approach at a high level -- i.e., add a new layer to the
network, called the {\it filter layer}, which ... }

@@

\textbf{Filter Layer: }

Filter layers have weights in the range $[0,+\infty]$ and are placed after
linear and convolutional layers of traditional deep networks \gl{This is how I
  used it but with the new implementation we can be more creative and put them
  anywhere, should we say "usually placed" insdead ?}. The \textit{Filter
  Layer} takes an input of size $\left(B \times C \times D_1 \times \dots
  \times D_n\right)$, where $B$ is the batch size, $C$ the number of features
(or channels), and $D$ any additional dimension. This structure makes it
compatible with fully connected layers with $n=0$ or convolutional layers with
$n=2$. Their crucial property is a parameter $\theta \in \mathbb{R}^C$, defined
as follows: \srm{Need to define B, C, D} \gl{=> done}.

@@

Where $\circ$ is the Hadamard product (pointwise multiplication) \gl{Should we
  keep only one of them to be more concise ?}, and $\theta$ is expanded in all
dimensions except the second one to match the input size. It is easy to see
that if for any $k$, if $\theta_k \leq 0$, the $k^{\text{th}}$ feature/channel
will forever be $0$. When this happens, we say the Filter layer disables the
neuron. These disabled neurons/channels can be removed from the network without
changing its output (we describe how we perform this removal below). We explain
next how the weights of the Filter Layer are adjusted during training.

@@

\srm{Give an English description of what this formalism
achieves -- what does the filter layer do, why is it significant.}

@@

This choice come from the Lasso loss [ref] and we use it for the same reason:
it introduces sparsity. The bigger the $\lambda$ the more entries in $\theta$
are set to zer, and since zeros entries in $\theta$ correspond to dead neurons,
lambda effectively control the number of neurons/channels in the entire
network. \srm{Again give some intuitive description of what is novel /
  important about this choice of loss function and learning process.  It's hard 
  to understand what is interesting/important about what you have done without
  you explicitly calling it out.} \gl{Is it better this way ?}

Next, we explain how to implement ShrinkNets efficiently.

\section{Software architecture}

\ra{this can be shorter. I'd add it as a third paragraph above, immediately
above the eval}

\gl{Here is a shorter (clearer ?) version:}

\par \textit{Filter Vectors} can easily be implemented in a few lines of code
in many existing deep learning frameworks. However, in ShrinkNets we assume
that we start with obviously oversized networks. If disabled neurons are not
quickly removed, the overhead might caused the training process to be
significantly slower than classic neural networks. This is why we want to
support what we call garbage collection: removing the weights in the layers
that are responsible(or using) features that are disabled.
\par To be able to catpure this relationship between layers we agumented
\textit{Pytorch} (the framework we used as the foundation of our implementation)
with a graph structure very similar to the one available in Keras [ref].
In this graph edges are effectively event hubs responsible for propagating
\textit{feature removal} events to subscribing layers. ShrinkNet layers are
agumented to emit and react to these events but also propagate the event
further if has an impact at the other endpoint of the layer (input or output).
\par This event-based implementation coordinated by edges makes it very easy
to integrate new layers in the library. We even provide automatic wrapping
for layers from pytorch if they have no internal state. For more complex ones,
we provide utilitaries that makes the implementation very concise.

@@

\gl{I rewrote the entire section}

@@

\ra{I'd remove this paragraph because we used it in the motivation already}
\par In the literature we can see different ways of approaching this issue. The
first one is to select parameters, train the model, evaluate it, repeat and pick
the best one. There are many different ways of selecting parameters: random
search [ref], grid search [ref], meta gradient descent [ref], Parsen trees[ref],
Gaussian processes [ref] etc... but they all suffer from one major drawback: we
have to train and evaluate a very large number of models; and even if some
algorithms allow parallel evaluations [ref] to reduce the overall training time,
the hardware and energy cost is still very significant.

\ra{I'd argue these techniques are aimed to improve inference time and not to
avoid hyper-parameter optimization}
\par To reduce the number of models trained, another approach is to train a
slightly bigger model and after convergence, remove as many parameters as
possible without impacting the performance of the model. Notable contributions
are Optimal Brain Damage [ref], Deep Compression [ref], and Group sparsity [ref]
(this one is especially related to this article). These techniques are very
interesting but they still require a reasonable network size to start with, so
they usually have to be combined with classic hyper-optimization techniques.

\srm{How is this shrinking
approach different than previous approaches like Optimal Brain Damage, Deep
Compression,e tc?}

\par However, some recent contributions like Non-Parametric Neural Netowrks
[ref] and [ref] try to learn the network size (width for the former and depth
for the latter) during the training process and without any prior assumption
about the network size. \ra{revise next}This method dynamically grows and shrinks the network
over time. Though this seems to be an attractive property to have, during our
experiments and according to their results, models are very slow to train and
sometimes converge to suboptimal solutions. Their method also requires a new
optimizer \textit{AdaRad}. \ra{explain how difficult is to get the optimizer
stable. that's a drawback of the method} \ra{Instead, ShrinkNets work out of the
box with existing libraries and does not need any changes to the optimizer.}

\srm{Give a short sentence about how our approach is
different than these previous approaches.}

\section{Future work}

\ra{this can be shorted, and used as conclusions}

\par Even though the firsts results this technique yields seem promising, there
are many area that we could explore to improve it. In the current implementation
we only "learn" the number of features (neurons or channels). We could try to
augment it with dynamic number of layers as seen in [ref] to be able to
determine the entire architecture.  \par We saw on Figure \ref{convergence_plot}
that the loss temporarily suffers from the removal of neurons. It is likely that
the loss would be more stable if the number of neurons converged faster or
neurons disappeared slower. For this reason we plan to explore proximal gradient
methods to optimize the filter vectors and/or randomize neuron removals.  \par
During our evaluation we picked small datasets mainly to be able to train many
models and have statistically significant distributions. With more computation
resources and time, we could see if it generalizes to bigger datasets and other
architectures like \texttt{ResNet} [ref] (small modifications to the existing
code base are required to support them)

\section{Conclusion}
Do the conclusion at the end

>>>>

@@

@@

\textbf{Filter Layers: } They have weights in the range $[0,+\infty]$ and are placed after
linear and convolutional layers of traditional deep networks \gl{This is how I
  used it but with the new implementation we can be more creative and put them
  anywhere, should we say "usually placed" insdead ?}. The \textit{Filter
  Layer} takes an input of size $\left(B \times C \times D_1 \times \dots
  \times D_n\right)$, where $B$ is the batch size, $C$ the number of features
(or channels), and $D$ any additional dimension. This structure makes it
compatible with fully connected layers with $n=0$ or convolutional layers with
$n=2$. Their crucial property is a parameter $\theta \in \mathbb{R}^C$, defined
as follows:

@@

Where $\circ$ is the Hadamard product (pointwise multiplication \gl{Should we
  keep only one of them to be more concise ?}), and $\theta$ is expanded in all
dimensions except the second one to match the input size. It is easy to see
that if for any $k$, if $\theta_k \leq 0$, the $k^{\text{th}}$ feature/channel
will forever be $0$. When this happens, we say the Filter layer disables the
neuron. These disabled neurons/channels can be removed from the network without
changing its output (we describe how we perform this removal below). We explain
next how the weights of the Filter Layer are adjusted during training.

@@

@@

This choice come from the Lasso loss [ref] and we use it for the same reason:
it introduces sparsity. The bigger the $\lambda$ the more entries in $\theta$
are set to zer, and since zeros entries in $\theta$ correspond to dead neurons,
lambda effectively control the number of neurons/channels in the entire
network.
Next, we explain how to implement ShrinkNets efficiently.

\textbf{Software architecture\footnote{The code is available as Python/PyTorch library on \url{http://github.com/mitdbg/fastdeepnets} \gl{Should we rename the repository ?}} :}\textit{Filter Vectors} can easily be implemented in a few lines of code
in many existing deep learning frameworks. However, in ShrinkNets we assume
that we start with obviously oversized networks. If disabled neurons are not
quickly removed, the overhead might caused the training process to be
significantly slower than classic neural networks. This is why we want to
support what we call garbage collection: removing the weights in the layers
that are responsible(or using) features that are disabled.
\par To be able to catpure this relationship between layers we agumented
\textit{Pytorch} (the framework we used as the foundation of our implementation)
with a graph structure very similar to the one available in Keras [ref].
In this graph edges are effectively event hubs responsible for propagating
\textit{feature removal} events to subscribing layers. ShrinkNet layers are
agumented to emit and react to these events but also propagate the event
further if has an impact at the other endpoint of the layer (input or output).
\par This event-based implementation coordinated by edges makes it very easy
to integrate new layers in the library. We even provide automatic wrapping
for layers from pytorch if they have no internal state. For more complex ones,
we provide utilitaries that makes the implementation very concise.

@@

@@

\par There are many methods in the litterature that try to simplify the neural
network. Most of them focuses on removing connections (eg: [ref], [ref]).
However, most hardware (GPU's, TPUS, TensorCores) do not really leverage sparsity
in the connections efficiently. On the other side, ShrinkNets and some others [ref]
and [ref] try to remove entire neuron instead of connections. This is very
usefull because it reduce the size of the matrices, producing speed-up on
every device. ShrinkNets improves on [ref] because it removes neurons during
training, speeding up the rest of the process and beats [ref] on convergence
speed and because it does not need to change the optimizer. To the best of our
knowledge this is also the first contribution that tries to learn the number of
channels of a convolutional neural network.

\section{Conclusion}

In this paper we presented a novel technique to guess a resonable network size
based on single parameter $\lambda$ that control the tradeof between loss and
the size. We demonstrated that it works both on fully connected networks and
convolutional neural networks.  Even though the firsts results seem promising,
there are many ways we could improve. In the current
implementation we only "learn" the number of features (neurons or channels). We
could try to augment it with dynamic number of layers as seen in [ref] to be
able to determine the entire architecture. \par We saw on Figure
\ref{convergence_plot} that the loss temporarily suffers from the removal of
neurons. It is likely that the loss would be more stable if the number of
neurons converged faster or neurons disappeared slower. For this reason we plan
to explore proximal gradient methods to optimize the filter vectors and/or
randomize neuron removals. \par During our evaluation we picked small datasets
mainly to be able to train many models and have statistically significant
distributions. With more computation resources and time, we could see if it
generalizes to bigger datasets and other architectures like \texttt{ResNet}
[ref] (small modifications to the existing code base are required to support
them)
